Ephemeral hierarchical data-management for neuroimaging applications

### 2023-06-28 paper/sea-neuro/paper-neuro.tex
Minor updates to text

\title{Automated neuroimaging data-management with Sea}

@@

	Neuroimaging open-data initiatives have led to increased availability of large
  scientific datasets. Whereas these datasets are shifting the processing
  bottleneck from compute-intensive to data-intensive, current standardized
  analysis tools lag behind in the adoption of strategies to mitigate the costs
  associated with data transfers. A major challenge in adapting neuroimaging
  applications for data-intensive processing is that they must be entirely
  rewritten.

  To facilitate data management for standardized neuroimaging tools, we
  developed Sea, a library that intercepts and redirects application read and
  write calls to minimize data transfer time. In this paper, we investigate the
  performance of Sea on three preprocessing pipelines implemented using standard toolboxes (FSL,
  SPM and AFNI), using three neuroimaging datasets of different sizes
  (OpenNeuro's ds001545, PREVENT-AD and the HCP dataset) on two high-performance
  computing clusters. Our results demonstrate that given the features of the
  applications and datasets, Sea provides large speedups (up to 32$\times$) when
  the shared file system's (e.g. Lustre) performance is deteriorated. When the
  shared file system is not overburdened by other users, performance is
  unaffected by Sea, suggesting that Sea's overhead is minimal even in cases
  where it is not useful. Overall, Sea is beneficial, even when performance gain
  is minimal, as it can be used to limit files on PFS.
\end{abstract}
\begin{document}

@@

    The recent explosion in publicly available neuroimaging data has lead to new data
    management challenges, from storage infrastructure application processing times \TG{missing word?}. To meet the storage, accessibility and
    security demands of neuroimaging data, large datasets have been stored in object stores provided by cloud storage providers \TG{mentioning clouds so early in the paper may give the impression that the paper will focus on them.}.
    Standardized metadata formats, such as the Brain Imaging Data Standards (BIDS)\cite{bids}, have been implemented to facilitate the sharing
    of the datasets. Tools such as DataLad~\cite{datalad} have been developed to provide versioning and provenance capture of data.
    Furthermore, recent developments in neuroimaging pipelines have addressed computation time limitations by adopting machine-learning approaches \TG{cite fastsurfer, hd-bet}.
    While all these solutions to Big Data-related data management exist, certain aspects, such as processing-related data-transfer overheads
    have received limited attention.

    The current largest neuroimaging datasets,
    the Human Connectome Project (HCP)~\cite{HCP} and the UK
    Biobank~\cite{ukbiobank}, reach up to Petabytes of data. For datasets of this
    volume, parallel high-capacity remote storage solutions are commonly relied
    upon as the persistent storage layer during processing. 
    Despite their respective high-scalability, these remote file systems cannot typically
    compete with the performance of compute-local storage (e.g. DRAM and SSDs). Should the pipelines 
    need to transfer large amounts of data to these remote storage locations, significant data
    processing overheads may occur.

    Big Data in neuroimaging can be present in two formats: 1) very large files
    (e.g. BigBrain~\cite{amunts2013bigbrain}) and 2) small files pertaining to a
    population of individuals, such as those found in MRI datasets, that
    collectively amount to very large datasets. Differences in the number and
    size of files result in different data transfer overheads. Whereas large
    file I/O is predominantly impacted by the underlying storage bandwidth,
    small file I/O results in a non-negligible latency overhead.

@@

    Newer neuroimaging applications leveraging popular neuroimaging pipeline engines
    also do not benefit from processing-related data management.
    Although engines such as Nipype~\cite{nipype} and
    Joblib~\cite{joblib} \TG{joblib is multithreading, dask or ray. I wouldn't call it an engine. You should use another example.} do not prohibit the use of data-management
    strategies, they do not facilitate the integration of these strategies into their resulting workflow. To give
    neuroimaging applications data management capabilities, the applications
    must interact with a file system or library that enable the strategies.

@@

    HPC systems rely on scalable network-based parallel file systems (e.g.
    Lustre) for storage. While such file systems offer excellent performance,
    they are shared between all the users on the cluster, meaning that users
    with data-intensive applications can effectively deteriorate the performance
    of the shared file system for all users on the cluster. Solutions for
    improving shared file system performance include
    throttling~\cite{huang2020ooops} the data-intensive workloads or
    recommending the use of Burst Buffers~\cite{bb} (e.g. reserving a compute
    node for storage or leveraging local compute storage during processing). The
    latter, however, requires that the user manages their data to and from the
    Burst Buffer, unless a Burst Buffer file system, such as
    BurstFS~\cite{burstfs} or GekkoFS~\cite{gekkofs} is installed.

    In order for a file system to be mounted by an HPC cluster user, it must be
    loadable without administrative privileges. As the applications are
    typically made to interact with POSIX-based file systems, the new file
    system must also be compliant to the format. One method to ensure that these
    conditions are met is by using the \texttt{LD\_PRELOAD} trick. This trick is
    used to intercept select library calls and redefine their behaviour. It has
    been used in many projects and to create lightweight userspace versions of
    file systems~\cite{xtreemfs,burstfs,gekkofs}.

    We present Sea~\cite{sea-comp}, a data-management library designed to reduce processing-related overheads. We view Sea as a complement
    to the existing stack of neuroimaging tools and standards meant to facilitate the processing of neuroimaging Big Data.    
    % \tristan{Intro logic is good but writing could be polished}
    % \tristan{Outlining the content of the paper would be useful}

@@

    \section{Materials and Methods}

@@

    \centering
    \includegraphics[width=\columnwidth]{figures/sea-diagram.pdf}%
\caption{The Sea data-management library overview}
\label{fig:seaneuro:diagram}
\end{figure*}
    Sea (Figure~\ref{fig:seaneuro:diagram}) is a data-management library that leverages the \texttt{LD\_PRELOAD}
    trick to intercept POSIX file system (more specifically, the GNU C library,
    glibc) calls on Linux systems. This enables Sea to redirect write calls
    aimed for a slower storage device to a faster device whenever possible.
    Similarly, when intercepting read calls, Sea can choose to read from a
    faster device if a copy is available on that device. Sea decides which
    storage location it can write to based on the detailed provided in an
    initialization file called \texttt{sea.ini}. This file informs Sea of which
    locations it can use to read and write to, and well as their order of
    priority.

@@

@@

    %Busy writers
    \section{Results}

\begin{subfigure}{\textwidth}
    \centering
    \captionsetup{width=.85\linewidth}
    \includegraphics[width=\columnwidth]{figures/slashbin_nozoom.pdf}%
    \caption{Complete results}\label{fig:seaneuro:slashbinfull}
\end{subfigure}
\begin{subfigure}{\textwidth}
    \centering
    \captionsetup{width=.85\linewidth}
    \includegraphics[width=\linewidth]{figures/slashbin_withzoom.pdf}
    \caption{Zoomed between 0 and 5000 seconds}\label{fig:seaneuro:slashbinzoom}
\end{subfigure}
\caption{Makespan comparison between Sea and Baseline on controlled dedicated cluster}
\label{fig:seaneuro:slashbin}
\end{figure*}

In Figure~\ref{fig:seaneuro:slashbin}, we observe that although performance is
variable, speedups are obtained by using Sea when Lustre
performance has been degraded by busy writers. On average, a single SPM
process preprocessing a single HCP image produced the greatest speedup
(13$\times$). AFNI preprocessing a single fMRI image of the PREVENT-AD dataset
using a single process was the next fastest pipeline, with an average speedup of
(5$\times$). FSL Feat had speedups as well, with a maximum average speedup of
1.3$\times$ when preprocessing a single PREVENT-AD image. SPM consistently had
excellent speedups, which is likely due to a mix of prefetching the initial
input files and the I/O patterns of the application. While it was expected that
AFNI would be the application that would benefit the most from Sea due to its
minimal compute time and the amount of data it generates, it did not benefit
from Sea to the extent that SPM did. As per Table~\ref{table:seaneuro:pipelines},
AFNI performs a very high number of glibc calls. While the overhead of an
individual call is likely very minimal, it is possible that the culmination of
all these calls results in a large enough bottleneck that affects both Sea and
Baseline executions an extensive amount. Moreover, the number of glibc calls
aimed at interacting with data on the Lustre storage are very few. AFNI still
obtained an average of 2$\times$ speedup across all experimental conditions.

\begin{figure*}
\centering
\includegraphics[width=\columnwidth]{figures/beluga_sea_tmpfs.pdf}%
\caption{Makespan comparison between Sea and tmpfs on the production cluster with flushing disabled}
\label{fig:seaneuro:tmpfs}
\end{figure*}

\begin{subfigure}{\textwidth}
    \centering
    \captionsetup{width=.85\linewidth}
    \includegraphics[width=\columnwidth]{figures/beluga_withoutflush_nozoom.pdf}%
    \caption{Complete results}\label{fig:seaneuro:belugafull}
\end{subfigure}
\begin{subfigure}{\textwidth}
    \centering
    \captionsetup{width=.85\linewidth}
    \includegraphics[width=\linewidth]{figures/beluga_withoutflush_zoom.pdf}
    \caption{Zoomed between 0 and 2000 seconds}\label{fig:seaneuro:belugazoom}
\end{subfigure}
\caption{Makespan comparison between Sea and Baseline on the production cluster with flushing disabled}
\label{fig:seaneuro:beluga-noflush}
\end{figure*}

\begin{figure*}
\begin{subfigure}{\textwidth}
    \centering
    \captionsetup{width=.85\linewidth}
    \includegraphics[width=\columnwidth]{figures/beluga_withflush_nozoom.pdf}%
    \caption{Complete results}\label{fig:seaneuro:belugafullwf}
\end{subfigure}
\begin{subfigure}{\textwidth}
    \centering
    \captionsetup{width=.85\linewidth}
    \includegraphics[width=\linewidth]{figures/beluga_withflush_zoom.pdf}
    \caption{Zoomed between 0 and 2000 seconds}\label{fig:seaneuro:belugazoomwf}
\end{subfigure}
\caption{Makespan comparison between Sea and Baseline on the production cluster when flushing on all files is enabled}
\label{fig:seaneuro:beluga-wflush}
\end{figure*}

    %Flushing - rely on modification time
    \section{Conclusion and Future work}

>>>>

\title{Ephemeral hierarchical data-management for neuroimaging applications}

@@

	Neuroimaging open-data initiatives have led to increased availability of large
  scientific datasets. While these datasets are shifting the processing
  bottleneck from compute-intensive to data-intensive, current standardized
  analysis tools have yet to adopt strategies that mitigate the costs
  associated with large data transfers. A major challenge in adapting neuroimaging
  applications for data-intensive processing is that they must be entirely
  rewritten.

  To facilitate data management for standardized neuroimaging tools, we
  developed Sea, a library that intercepts and redirects application read and
  write calls to minimize data transfer time. In this paper, we investigate the
  performance of Sea on three preprocessing pipelines implemented using standard toolboxes (FSL,
  SPM and AFNI), using three neuroimaging datasets of different sizes
  (OpenNeuro's ds001545, PREVENT-AD and the HCP dataset) on two high-performance
  computing clusters. Our results demonstrate that given the features of the
  applications and datasets, Sea provides large speedups (up to 32$\times$) when
  the shared file system's (e.g. Lustre) performance is deteriorated. When the
  shared file system is not overburdened by other users, performance is
  unaffected by Sea, suggesting that Sea's overhead is minimal even in cases
  where its benefits are limited. Overall, Sea is beneficial, even when performance gain
  is minimal, as it can be used to limit files on PFS.
\end{abstract}
\begin{document}

@@

    The ever-growing surge in publicly available neuroimaging data has lead to new data
    management challenges, from storage infrastructure to application processing times.
    Many tools have been developed by the community to address to various needs in 
    neuroimaging data management.
    %To meet the storage, accessibility and security demands of neuroimaging data,
    %large datasets have been stored in object stores provided by cloud storage providers
    %\TG{mentioning clouds so early in the paper may give the impression that the paper will 
    % focus on them.}.
    For instance, standardized metadata formats, such as the Brain Imaging Data Standards (BIDS)\cite{bids},
    have been implemented to facilitate the sharing
    of the datasets. Tools such as DataLad~\cite{datalad} have been developed to provide 
    versioning and provenance capture of data.
    Furthermore, recent developments in neuroimaging pipelines have addressed computation time 
    limitations by adopting machine-learning approaches \TG{cite fastsurfer, hd-bet}.
    While all these solutions to Big Data-related data management exist, certain aspects, 
    such as processing-related data-transfer overheads
    have received limited attention.

    The current largest neuroimaging datasets,
    the Human Connectome Project (HCP)~\cite{HCP} and the UK
    Biobank~\cite{ukbiobank}, reach up to Petabytes of data. For datasets of this
    volume, parallel high-capacity remote storage solutions are commonly relied
    upon as the persistent storage layer during processing. 
    Despite their respective high-scalability, these remote file systems cannot typically
    compete with the performance of compute-local storage (e.g., tmpfs and SSDs). Should the pipelines 
    need to transfer large amounts of data to these remote storage locations, significant data
    processing overheads may occur.

    Big Data in neuroimaging max exist in two formats: 1) very large files
    (e.g. BigBrain~\cite{amunts2013bigbrain}) and 2) small files pertaining to a
    population of individuals, such as those found in MRI datasets, that
    collectively amount to very large datasets. Differences in the number and
    size of files result in different data transfer overheads. Whereas large
    file I/O is predominantly impacted by the underlying storage bandwidth,
    small file I/O results in a non-negligible latency overhead.

@@

    Newer neuroimaging applications leveraging popular neuroimaging pipeline engines
    also do not benefit from processing-related data management.
    Although engines such as Nipype~\cite{nipype}
    % and Joblib~\cite{joblib} \TG{joblib is multithreading, dask or ray. I wouldn't call it an engine. You should use another example.}
    do not prohibit the use of data-management
    strategies, they do not facilitate the integration of these strategies into
    their resulting workflow. To give
    neuroimaging applications data management capabilities, the applications
    must interact with a file system or library that enable the strategies.

@@

    HPC systems rely on scalable network-based parallel file systems (e.g.,
    Lustre) for storage. While such file systems offer excellent performance,
    they are shared between all the users on the cluster, meaning that users
    with data-intensive applications can effectively deteriorate the performance
    of the shared file system for all users on the cluster. Solutions for
    improving shared file system performance include
    throttling~\cite{huang2020ooops} the data-intensive workloads or
    recommending the use of Burst Buffers~\cite{bb} (e.g., reserving a compute
    node for storage or leveraging local compute storage during processing). The
    latter, however, requires that the user manages their data to and from the
    Burst Buffer, unless a Burst Buffer file system, such as
    BurstFS~\cite{burstfs} or GekkoFS~\cite{gekkofs} is installed.

    In order for a file system to be mounted by an HPC cluster user, it must be
    loadable without administrative privileges. As the applications are
    typically made to interact with POSIX-based file systems, the new file
    system must also be compliant to the format. One method to ensure that these
    conditions are met is through the \texttt{LD\_PRELOAD} trick. This trick is
    used to intercept select library calls and redefine their behaviour. It has
    been used in many projects and to create lightweight userspace versions of
    file systems~\cite{xtreemfs,burstfs,gekkofs}.

    In this paper, we present Sea~\cite{sea-comp}, a data-management library designed to reduce
    transfer-related overheads. We view Sea as a complement
    to the existing stack of neuroimaging tools and standards meant to facilitate 
    the processing of neuroimaging Big Data.   
    % \tristan{Intro logic is good but writing could be polished}
    % \tristan{Outlining the content of the paper would be useful}

@@

    \section{Results}

@@

    \centering
    \includegraphics[width=\columnwidth]{figures/sea-diagram.pdf}%
\caption{The Sea data-management library overview}
\label{fig:seaneuro:diagram}
\end{figure*}
    Sea (Figure~\ref{fig:seaneuro:diagram}) is a data-management library that leverages
    the \texttt{LD\_PRELOAD}
    trick to intercept POSIX file system (more specifically, the GNU C library,
    glibc) calls on Linux systems. This enables Sea to redirect write calls
    aimed at slower storage devices to a faster devices whenever possible.
    Similarly, when intercepting read calls, Sea can choose to read from a
    faster device if a copy is available on that device. Sea decides which
    storage location it can write to based on the details provided in an
    initialization file called \texttt{sea.ini}. This file informs Sea of which
    locations it can use to read and write to, and well as their order of
    priority. It is important to note that Sea is not a filesystem, but rather a
    lightweight tool that redirects files to more optimal storage devices.

@@

\begin{subfigure}{0.5\textwidth}
    \centering
    \captionsetup{width=.85\linewidth}
    \includegraphics[width=\columnwidth]{figures/slashbin_nozoom.pdf}%
    \caption{Complete results}\label{fig:seaneuro:slashbinfull}
\end{subfigure}
\begin{subfigure}{0.5\textwidth}
    \centering
    \captionsetup{width=.85\linewidth}
    \includegraphics[width=\linewidth]{figures/slashbin_withzoom.pdf}
    \caption{Zoomed between 0 and 5000 seconds}\label{fig:seaneuro:slashbinzoom}
\end{subfigure}
\caption{Makespan comparison between Sea and Baseline on controlled dedicated cluster}
\label{fig:seaneuro:slashbin}
\end{figure*}

In Figure~\ref{fig:seaneuro:slashbin}, we observe that although performance is
variable, speedups are obtained by using Sea when Lustre
performance has been degraded by busy writers. On average, a single SPM
process preprocessing a single HCP image produced the greatest speedup
(13$\times$). AFNI preprocessing a single fMRI image of the PREVENT-AD dataset
using a single process was the next fastest pipeline, with an average speedup of
(5$\times$). FSL Feat had speedups as well, with a maximum average speedup of
1.3$\times$ when preprocessing a single PREVENT-AD image. SPM consistently had
excellent speedups, which is likely due to a mix of prefetching the initial
input files and the I/O patterns of the application. While it was expected that
AFNI would benefit the most from Sea due to its
minimal compute time and the amount of data it generates, it did not benefit
from Sea to the extent that SPM did. As per Table~\ref{table:seaneuro:pipelines},
AFNI performs a very high number of glibc calls. While the overhead of an
individual call is likely very minimal, it is possible that the culmination of
all these calls results in a large enough bottleneck that affects both Sea and
Baseline executions an extensive amount. Moreover, the number of glibc calls
aimed at interacting with data on the Lustre storage are very few. AFNI still
obtained an average of 2$\times$ speedup across all experimental conditions.

\begin{figure*}
\centering
\includegraphics[width=0.7\columnwidth]{figures/beluga_sea_tmpfs.pdf}%
\caption{Makespan comparison between Sea and tmpfs on the production cluster with flushing disabled}
\label{fig:seaneuro:tmpfs}
\end{figure*}

\begin{subfigure}{0.5\textwidth}
    \centering
    \captionsetup{width=.85\linewidth}
    \includegraphics[width=\columnwidth]{figures/beluga_withoutflush_nozoom.pdf}%
    \caption{Complete results}\label{fig:seaneuro:belugafull}
\end{subfigure}
\begin{subfigure}{0.5\textwidth}
    \centering
    \captionsetup{width=.85\linewidth}
    \includegraphics[width=\linewidth]{figures/beluga_withoutflush_zoom.pdf}
    \caption{Zoomed between 0 and 2000 seconds}\label{fig:seaneuro:belugazoom}
\end{subfigure}
\caption{Makespan comparison between Sea and Baseline on the production cluster with flushing disabled}
\label{fig:seaneuro:beluga-noflush}
\end{figure*}

\begin{figure*}
\begin{subfigure}{0.5\textwidth}
    \centering
    \captionsetup{width=.85\linewidth}
    \includegraphics[width=\columnwidth]{figures/beluga_withflush_nozoom.pdf}%
    \caption{Complete results}\label{fig:seaneuro:belugafullwf}
\end{subfigure}
\begin{subfigure}{0.5\textwidth}
    \centering
    \captionsetup{width=.85\linewidth}
    \includegraphics[width=\linewidth]{figures/beluga_withflush_zoom.pdf}
    \caption{Zoomed between 0 and 2000 seconds}\label{fig:seaneuro:belugazoomwf}
\end{subfigure}
\caption{Makespan comparison between Sea and Baseline on the production cluster when flushing on all files is enabled}
\label{fig:seaneuro:beluga-wflush}
\end{figure*}

    %Flushing - rely on modification time
    \section{Materials and Methods}

@@

    %Busy writers
    \section{Conclusion and Future work}
